### 2014-07-26 doc/thesis/introduction.tex
Starting the rewriting of the Typed Lua chapter

In this work we present the design and implementation of Typed Lua:
an optional type system for Lua that is complex enough to
preserve some of the Lua idioms that programmers are already used to,
but that also includes new constructs that help programmers
structure Lua programs.

>>>>

In this work we present the design and implementation of Typed Lua:
an optional type system for Lua that is rich enough to
preserve some of the Lua idioms that programmers are already used to,
but that also includes new constructs that help programmers
structure Lua programs.

### 2014-07-26 doc/thesis/typedlua.tex
Starting the rewriting of the Typed Lua chapter

\section{Typed Lua through examples}

Our main goal is to allow programmers to combine static and dynamic
typing in Lua, so we propose Typed Lua as an extension to Lua 5.2
syntax that introduces optional type annotations, and class-based
object-oriented programming through the definition of classes,
interfaces, and modules.
An important requirement of Typed Lua is that it should be backwards
compatible with Lua, that is, it should allow programmers to run any
Lua code without modifications.
For this reason, these syntactic extensions include new keywords
that are not reserved words.

Figure \ref{fig:typelang} shows that Typed Lua includes first-class
and second-class types.
Typed Lua uses first-class types for representing Lua values and
second-class types for defining the signatures of functions.
First-class types include literal types, base types, the type $\Any$,
union types, function types, table types, and recursive types.
Second-class types include the type $\Void$, first-class types,
vararg types, and tuple types.
Literal types are types that represent a single value
(e.g., $\Nil$, $\False$, $\True$, literal numbers, and
literal strings).
Literal types are subtypes of their respective base types.
More precisely, the literals $\False$ and $\True$ are subtypes of
$\Boolean$, any literal number $n$ is subtype of $\Number$, and any
literal string $s$ is subtype of $\String$.
The type $\Any$ is the dynamic type that allows dynamic code coexist
with static code.
We will discuss union types, function types, table types, recursive
types, and second-class types along this section.

\begin{figure}[!ht]
\textbf{Type Language}\\
\dstart
$$
\begin{array}{rlr}
T ::= & & \emph{first-class types}\\
& \;\; L & \emph{literal types}\\
& | \; B & \emph{base types}\\
& | \; \Any & \emph{dynamic type}\\
& | \; T \cup T & \emph{union type}\\
& | \; S \rightarrow S & \emph{function type}\\
& | \; \{T:T, ..., T:T\} & \emph{table type}\\
& | \; X & \emph{type variable}\\
& | \; \mu X.T & \emph{recursive type}\\
L ::= & \Nil \; | \; \False \; | \; \True \; | \;
  {<}{\it number}{>} \; | \; {<}{\it string}{>} & \\
B ::= & \Boolean \; | \; \Number \; | \; \String & \\
S ::= & \Void \; | \; V & \emph{second-class types}\\
V ::= & \;\; T & \\
& | \; T* & \emph{vararg type}\\
& | \; T \times V & \emph{tuple type}
\end{array}
$$
\dend
\caption{Typed Lua types}
\label{fig:typelang}
\end{figure}

@@

@@

@@

>>>>

\section{Typed Lua}

The main goal of Typed Lua is to allow programmers to combine static
and dynamic typing in Lua, so Typed Lua is an extension to Lua 5.2
syntax that introduces optional type annotations, and class-based
object-oriented programming through the definition of classes,
interfaces, and modules.
An important feature of Typed Lua is its backwards compatibility with
Lua, that is, unannotated Typed Lua code is valid Lua code.
For this reason, the syntactic extensions introduced by Typed Lua
include new keywords that are not reserved words.
Appendix \ref{app:syntax} presents the complete syntax of Typed Lua.

\subsection{Type Language}

Lua values can have one of eight tags:
\emph{nil}, \emph{boolean}, \emph{number}, \emph{string},
\emph{function}, \emph{table}, \emph{userdata}, and \emph{thread}.
In this section we present the types that Typed Lua introduces
to represent these tags.

\begin{figure}[!ht]
\textbf{Type Language}\\
\dstart
$$
\begin{array}{rlr}
\multicolumn{3}{c}{\textbf{First-level types}}\\
T ::= & \;\; L & \textit{literal types}\\
& | \; B & \textit{base types}\\
& | \; \Nil & \textit{nil type}\\
& | \; \Value & \textit{top type}\\
& | \; \Any & \textit{dynamic type}\\
& | \; \Self & \textit{self type}\\
& | \; T \cup T & \textit{disjoint union types}\\
& | \; P \rightarrow R & \textit{function types}\\
& | \; \{ \}_{o|c} & \textit{empty table type}\\
& | \; \{F, ..., F\}_{o|c} & \textit{table types}\\
& | \; X & \textit{type variables}\\
& | \; \mu X.T & \textit{recursive types}\\
& | \; X_{i} & \textit{projection types}\\
L ::= & \False \; | \; \True \; | \; {<}{\it number}{>} \; | \; {<}{\it string}{>} & \\
B ::= & \Boolean \; | \; \Number \; | \; \String & \\
F ::= & K:T \; | \; \Const \; K:T & \textit{field types}\\ 
K ::= & L \; | \; B \; | \; \Value \; | \; \Any & \textit{key types}\\
\multicolumn{3}{c}{}\\
\multicolumn{3}{c}{\textbf{Second-level types}}\\
S ::= & \;\; T &\\
& | \; T* & \textit{vararg types}\\
& | \; T \times S & \textit{tuple types}\\
P ::= & \Void \; | \; S & \textit{parameters list type}\\
R ::= & \Void \; | \; S \; | \; S \cup R & \textit{return types}\\
\end{array}
$$
\dend
\caption{Typed Lua types}
\label{fig:typelang}
\end{figure}

Figure \ref{fig:typelang} presents the abstract syntax of the
Typed Lua types.
We split Typed Lua types into two categories:
\emph{first-level types} and \emph{second-level types}.
We use first-level types to represent Lua values, and
second-level types to type expression lists, multiple assignments,
and function applications.
We will all Typed Lua types along this section.

\subsection{Subtyping and Consistent-Subtyping}

\subsection{Atomic types and Functions}

@@

\subsection{Unions}

@@

\subsection{Tables and Interfaces}

@@

\subsection{Modules}

\subsection{Objects and Classes}
